---
title: "Intro to R for Seurat"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing packages in r

Installation of R Packages takes place within R using R scripts. The syntax for installing packages is as follows:

  install.packages("package_name")

Packages are installed once per the computer you are using, though if you need to update a package, you would want to rerun install.packages().

After packages are installed (once per machine), packages need to be loaded into your project everytime a project's global environment has been reset. Packages are loaded using the library function:

  library(package_name)

```{r}
# install and load the seurat package
install.packages('Seurat', repos = "http://cran.us.r-project.org")
library(Seurat)

# install and load the dpylr package
install.packages("dplyr", repos = "http://cran.us.r-project.org")
library(dplyr)
```

## Creating variables in R

Variables allow you to store information within keywords. This stored information can be a single data element, like a string, a number, or a boolean, or variables can be used to store more complicated data structures such as lists, dataframes, graphs, and custom user-created functions. 

Variables in R are created using the following notation: <-

#### Single datum and lists stored as variables
```{r cars}
x <- 1
x
numerical_list = c(0.75, 10, 8, 7, 5, 6.5)
numerical_list
list_of_strings <- c("Elo", "Charlie", "Molly", "Woody", "Monte", "Ben")
list_of_strings
list_of_booleans <- c(TRUE, FALSE, FALSE, FALSE, TRUE, FALSE)
list_of_booleans
```

#### Data structures stored as variables

You can also embed plots, for example:

```{r pressure, echo=FALSE}
df <- data.frame(
  list_of_strings,
  numerical_list,
  list_of_booleans
)
df
```

## Using functions in R

Functions allow you to perform complicated tasks in R without custom codeing every detail. Functions are either part of Base R, or they are tied to a particular package. Base R includes mathmatical, basic stats and basic graphs. Packages add additional functions that may allow for more customization or more specific analysis types.   

Functions take arguments and provide results.

#### Mathmatical functions in Base R
```{r}
mean(df$numerical_list)
sd(df$numerical_list)
```

#### Vector functions in Base R
```{r}
sort(df$list_of_strings)
```
#### Referencing documentation about functions

The help() fucntion and the ? help operator in R provide access to R documentation. 

Referencing documentation for the c list creating function:

```{r}
?c
```

Referencing documentation for the averaging function, specifying that you would like the mean function from base R (opposed to another package which might also have a mean function):

```{r}
?base::mean
```

## 